Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-03-17 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add HGmodel manuscript in ACS format

% \documentclass[rmp,aps,floatfix,authordate1-4,preprint]{revtex4}
% \documentclass[prb,aps,floatfix,authordate1-4,preprint]{revtex4}

\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[obeyFinal]{easy-todo}
%\usepackage{scrtime}
\usepackage[multiple]{footmisc}
%\usepackage[sort&compress]{natbib}
%\bibpunct{(}{)}{,}{n}{}{}
%\renewcommand{\bibnumfmt}[1]{#1.}
\usepackage[normalem]{ulem}

\begin{document}
%\singlespacing

\onecolumngrid
{\color{red}
Matti Javanainen has made the text modifications to the manuscript in excellent way. He modified directly the manuscript 
tex file available in the GitHub and then made a pull request to add a new version of the manuscript into the GitHub repository. 
In addition, he used latexdiff which can be generated to show the pdf with changes.

After Matti had added his version I looked the changes through and agreed. Then I changed his version to the up-to-date version 
and made some minor modifications.

All the mentioned files can be found from the GitHub repository
https://github.com/NMRLipids/nmrlipids.blogspot.fi

up-to date manuscript
HGmodel\_draft3.tex

Version Matti uploaded
HGmodel\_draft3\_matti.tex

pdf showing the changes made by matti:
HGmodel\_draft3\_matti\_latexdiff.pdf

I think that this is a very good and convenient way to work collaboratively with the manuscript. If you want you can also make 
a pull request directly using the up-to-date manuscript file name (HGmodel\_draft3.tex). Then I will accept the pull request, 
or we will do some discussion before acception. Using latexdiff is very good, but the history will also be saved automatically 
to the GitHub.

If you find this difficult (as it may be if you are not familiar with GitHub) you can also send your comments and modifications 
in some other way. However, keep in mind that it is very important to clearly report somehow what you have done. 
This is to ease my work to handle the input from large amount of contributors.
}

\title{Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions}

\author{Alexandru Botan}
\thanks{The authors are listed in alphaphetical order.}
\thanks{The author list is not completed.}
\thanks{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\thanks{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela}
\author{Patrick Fuchs}
\thanks{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
%\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
\author{Matti Javanainen}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Waldemar Kulig}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\thanks{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
%\author{Alexander Lyubartsev}
\author{Markus S. Miettinen}
\thanks{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
\author{Luca Monticelli}
\thanks{5IBCP, CNRS UMR 5086, Lyon, France}
\author{Jukka M\"a\"att\"a}
\thanks{Aalto University, Espoo, Finland}
%\setcounter{page}{1}
%\author{Vasily S. Oganesyan}
%\thanks{University of East Anglia, Norwich, United Kingdom}
\author{O. H. Samuli Ollila} 
\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
\author{Marius Retegan}
\thanks{Max Planck Institute for Chemical Energy Conversion, M\"ulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\thanks{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
\thanks{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Joona Tynkkynen}
\thanks{Tampere University of Technology, Tampere, Finland}
%\author{Mark Wilson}
%\author{Alexander Vogel}

\begin{abstract}
Phospholipids are essential building blocks of biological membranes.
Despite of vast amount of accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations model 
would automatically resolve the structures giving an interpretation of experimental results, if the model
reproduced the experimental data. In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
atomistic models, and we compare simulations with experiments (in fully hydrated conditions) in terms of C--H bond vector order
parameters for the glycerol backbone and choline headgroups. None of the current models is not accurately enough to resolve the structure with experimental accuracy.
However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
would resolve the structure. Despite of the inaccuracy in the fully hydrated
structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
between lipids and cholesterol better than the Berger/H\"oltje model.
{\bf This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
communication platform. All the scientific contributions have been done through the blog and
are publicly available. In addition, almost all simulation trajectories and files are made available
in the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids} which has became the most extensive
publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.
}
%This work has been, and continues to be, progressed and discussed through the blog: \url{nmrlipids.blogspot.fi}. 
%Everyone is invited to join the discussion and make contributions through the blog. 
%The manuscript will be eventually submitted to an appropriate scientific journal. 
%Everyone who has contributed to the work through the blog will be offered 
%coauthorship. For more details see: \url{nmrlipids.blogspot.fi}.

\end{abstract}

\maketitle

~\vspace{0.3cm}\\
{\it \bf}

@@

Phospholipids containing various polar headgroups and acyl chains are essential building blocks of 
biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
and theoretical techniques as a simple model for cellular membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
Phospholipid molecules are composed of hydrophobic acyl chains which are connected by a glycerol backbone to a hydrophilic headgroup.
See Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
The behaviour of the acyl chains in a lipid bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
The conformations sampled by the glycerol backbone and choline in a fluid bilayer are, however, not fully 
resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
techniques give only a set of values that the structure has to fulfil, but
there is no unique way to derive the actual structure from them~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
Some structural details have been extracted from crystal structure, $^1$H NMR studies and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06}
but general consensus about the structures sampled in the fluid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}. 
On the other hand, the glycerol backbone structures are similar for various biologically
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments~\cite{gally81} and the headgroup choline structures are similar in model membranes and
real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
Thus, the solution of phosphatidylcholine glycerol backbone and choline structures would be 
useful for understanding a wide range of different biological membranes.

@@

In this work we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12} and MacRog models~\cite{maciejewski14}
have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
structures between the most often used (Berger based) lipid model~\cite{berger97} 
\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.
%Based on the results and discussions with Olle Edholm and Peter Tieleman in the blog, it seems that 
%the glycerol backbone dihedrals are not the same in the model used in the publication by Berger et al.~\cite{berger97}
%and in the model shared in the webpage of Peter Tieleman (http://wcm.ucalgary.ca/tieleman/downloads).
%In this work (as it is quite often done in other works as well) we currently refer to the model in the webpage as ''Berger'' model.
%This is misleading. We have to call it something else. One suggestion could be to call it ''BergerM'', because it was firt time
%used in Marrink et al. Biophys. J. 74,931-943 (1998). Any other suggestions?\\
%COMMENT: Patrick Fuchs suggested ''Berger-T''. This is better. \\
%COMMENT2: Current trial, based in blog discussion, is to use Berger-MOLECULE-YEAR [citation].
}
and the best performing models, to demonstrate that by using the 
order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
is accurate enough to properly resolve the atomistic resolution stuctures.

@@

\subsection{Open collaboration}
\todo{Should we write more about this?}

This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi} as a communication platform.
The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
{\bf The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
in a Berger based model (the most used molecular dynamics simulation model for lipid bilayers).
After publishing the initial manuscript, the open invitation for further contributions and discussion through the blog was presented.
All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship. 
The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
The contributions from each author are summarized in ?? (the location depends on the journal probably).

Almost all simulation data, including running parameters for reprocution and trajectories for further analysis, are collected
to the nmrlipids community in Zenodo file repository \url{https://zenodo.org/collection/user-nmrlipids}.
Thus, besides the main topic of this manuscript we present the most extensive available collection of simulations trajectories
for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub organization \url{https://github.com/NMRlipids}.
}

\subsection{Order parameters from experiments}\label{expORDp}
\todo{DONE
%We should mention the sign issue and refer to the blog post:
%http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
%Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
%and simulations" or similar, and also move the "forking" explanation here.
}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.
COMMENT: There is also the option to put the most relevant blog posts as a supplementary material. The advantage of this would
be that the posts would be peer reviewed as well.}.

{\bf The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
However, only $^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
For more detailed discussion about the sign measurements of the order parameters see \url{http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\author{Andrew N. Other}
%\altaffiliation{A shared footnote}
%\author{Fred T. Secondauthor}
%\altaffiliation{Current address: Some other place, Othert\"own,
%Germany}
%\author{I. Ken Groupleader}
%\altaffiliation{A shared footnote}
%\email{i.k.groupleader@unknown.uu}
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}
%\author{Susanne K. Laborator}
%\email{s.k.laborator@bigpharma.co}
%\affiliation[BigPharma]
%{Lead Discovery, BigPharma, Big Town, USA}
%\author{Kay T. Finally}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}

\author{Alexandru Botan}
%\affiliation{The authors are listed in alphaphetical order.}
%\affiliation{The author list is not completed.}
\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
%\author{Andrea Catte}
%\affiliation{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela}
\affiliation[Mexico]{Mexico}
\author{Patrick Fuchs}
\affiliation[CNRS paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\alsoaffiliation[Diderot University]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Matti Javanainen}
\affiliation[Tampere University of Technology]
{Tampere University of Technology}
\author{Waldemar Kulig}
\affiliation[Tampere University of Technology]
{Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\affiliation[Kyoto University]
{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
\author{Markus S. Miettinen}
\affiliation[Freie Universitat Berlin] 
{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
\author{Luca Monticelli}
\affiliation[UMR] 
{5IBCP, CNRS UMR 5086, Lyon, France}
\author{Jukka Maatta}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@aalto.fi.}
\affiliation[Aalto University]
{Aalto University, Espoo, Finland}
\author{Marius Retegan}
\affiliation[Max Planck]
{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
\author{Tomasz Rog}
\affiliation[Tampere University of Technology]
{Tampere University of Technology, Tampere, Finland}
\author{Hubert Santuz}
\affiliation[INSERM]
{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
\alsoaffiliation[Diderot]
{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
\author{Joona Tynkkynen}
\affiliation[Tampere University of Technology]
{Tampere University of Technology, Tampere, Finland}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Phospholipids are essential building blocks of biological membranes.
Despite of vast amount of accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations model 
would automatically resolve the structures giving an interpretation of experimental results, if the model
reproduced the experimental data. 
In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
atomistic models, and we compare simulations with experiments (in fully hydrated conditions) in terms of C--H bond vector order
parameters for the glycerol backbone and choline headgroups. 
We show that this comparison can be used to judge the atomistic resolution structural accuracy of the models,
thus allowing the usage of molecular dynamics simulations to interpret the structures of biomolecules in 
biologically relevant conditions from NMR experiments. Further, we review previous experimental
data for dehydrated lipid bilayers and cholesterol containing bilayers, and interpretate it with simulations.
The interpretation reveals that the increased choline order parameters indicate P--N vector tilting more parallel to membrane,
and that cholesterol induces only minor changes to the glycerol backbone structure.
%None of the current models is not accurately enough to resolve the structure with experimental accuracy.
%However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
%that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
%would resolve the structure. Despite of the inaccuracy in the fully hydrated
%structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
%is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
%between lipids and cholesterol better than the Berger/H\"oltje model.
This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
communication platform. All the scientific contributions have been done through the blog and
are publicly available. In addition, almost all simulation trajectories and files are made available
in the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids} which has became the most extensive
publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Phospholipids containing various polar headgroups and acyl chains are essential building blocks of 
biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
and theoretical techniques as a simple model for cellular membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
Phospholipid molecules are composed of hydrophobic acyl chains which are connected by a glycerol backbone to a hydrophilic headgroup.
See Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
The behaviour of the acyl chains in a lipid bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
The conformations sampled by the glycerol backbone and choline in a fluid bilayer are, however, not fully 
resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
techniques give only a set of values that the structure has to fulfil, but
there is no unique way to derive the actual structure from them~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
Some structural details have been extracted from crystal structure, $^1$H NMR studies and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06}
but general consensus about the structures sampled in the fluid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}. 
On the other hand, the structural parameters for the glycerol backbone are similar for various biologically
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments~\cite{gally81} and the structural parameters for the headgroup choline structures are similar in model membranes and
real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
Thus, the solution of phosphatidylcholine glycerol backbone and choline structures would be 
useful for understanding a wide range of different biological membranes.

@@

In this work we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12} and MacRog models~\cite{maciejewski14}
have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
structures between the most often used (Berger based) lipid model~\cite{berger97} 
%\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.}
and the best performing models, to demonstrate that by using the 
order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
is accurate enough to properly resolve the atomistic resolution stuctures.

@@

\subsection{Open collaboration}
%\todo{Should we write more about this?}

This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}
\todo{Current idea is that we will assign doi for the blog by using the help from  winnower administrators:
https://thewinnower.com/posts/archiving-and-aggregating-alternative-scholarly-content-dois-for-blogs.
Once done, that doi will be cited here and other places where blog content needs to be cited.} 
as a communication platform.
The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
in a Berger based model (the most used molecular dynamics simulation model for lipid bilayers).
After publishing the initial manuscript, the open invitation for further contributions and discussion through the blog was presented.
All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship
according to the guidelines defined in the beginning of the project\todo{Citation to the On Credits in the blog here}. 
The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
The contributions from each author are summarized in the Supplementary Information. %(the location depends on the journal probably).

Almost all simulation data, including running parameters for reprocution and trajectories for further analysis, are collected
to the nmrlipids community in Zenodo file repository \url{https://zenodo.org/collection/user-nmrlipids}.
Thus, besides the main topic of this manuscript we present the most extensive available collection of simulations trajectories
for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
organization \url{https://github.com/NMRlipids} \todo{Doi will be assigned to the GitHub repository before submission}.

\subsection{Order parameters from experiments}\label{expORDp}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
\todo{Citation to the blog will be added when available.}

The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
However, only $^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
For more detailed discussion about the sign measurements of the order parameters see \url{http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html}
\todo{Citation to the blog will be added when available.}

For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
can be observed with both $^2$H NMR and $^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
separate populations of lipid conformations~\cite{engel81,gally81}.
